%% ============================================================
%%  ABSTRACT
%%  v5.0.0rc2 — Updated June 2026
%% ============================================================
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
\thispagestyle{plain}

Software vulnerabilities remain a critical challenge in modern development. Traditional static application security testing (SAST) tools suffer from fragmented output formats, high false-positive rates, and an inability to reason about context-specific risks. Integrating multiple complementary tools adds significant engineering overhead, and the absence of AI-assisted interpretation leaves developers overwhelmed by uncontextualised findings.

This project presents \textbf{\ACRQA{}} (Automated Code Review \& Quality Assurance), a self-hosted, AI-augmented SAST platform that orchestrates nineteen open-source engines across Python, JavaScript/TypeScript, and Go. Heterogeneous outputs are normalised into a single \texttt{CanonicalFinding} schema and enriched through a Retrieval-Augmented Generation (RAG) pipeline backed by a curated knowledge base of 66 security rules and Groq LLaMA-3.3-70B. The twelve-stage pipeline incorporates parallel tool execution, cross-tool deduplication, intra-procedural taint analysis, AST call-graph reachability gating, five-signal label-free confidence scoring, RAG enrichment with N-gram entropy consistency filtering, a NIST PQC-aware Cryptographic Bill of Materials, quality gates, and dual-signature attestation (ECDSA-P256 + CRYSTALS-Dilithium3). The full stack ships as an 8-service Docker Compose deployment with a 52-endpoint FastAPI service and a WCAG~2.1~AA React~18 + TypeScript dashboard at zero recurring licensing cost.

Evaluation spans four corpora: (i)~Layer~A ground truth (DVPWA, Pygoat, VulPy, DSVW) --- \textbf{97.1\% overall} and \textbf{100\% security-class precision} on 836 findings; (ii)~Layer~B 30-repo adversarial corpus of mature production libraries --- the \textbf{P4 Confirmed Tier} (HIGH severity + 22 curated rules + production path + Bandit-internal HIGH confidence) achieves \textbf{96.4\% conservative / 100\% optimistic precision} (95\% CI [90.9\%, 100\%]), F\textsubscript{1}~98.2\%, clearing the $\geq$80\% industry threshold for autonomous PR-blocking; (iii)~20-CVE recall battery --- \textbf{100\% recall on detectable CVEs (8/8)}; (iv)~four novel evaluations --- live-CVE blind holdout (X1), Docker-sandbox exploit verification across \textbf{13 exploit categories} (X2), 400-sample AI-generated code study (X3, finding 8--12$\times$ the human baseline), and a time-aware predictive backtest on Django (X4). A unified head-to-head against Bandit (F\textsubscript{1}~21.8\%) and Semgrep CE (45.7\%) on the same corpora confirms \ACRQA{} as the only evaluated tool with both 100\% CVE recall and near-perfect precision in its high-confidence stratum. On the RealVuln 2026 real-world leaderboard (1,000 production CVEs), \ACRQA{}'s anchoring technique matches Semgrep's 17.5\% and Snyk's 17.4\% detection rate while SonarQube reaches 6.5\%, corroborating that the Confirmed Tier precision gap reflects category difficulty, not tool deficiency. The 2,897-test suite reaches 84.89\% CORE coverage; the REST layer sustains 500 RPS at $<$2\,s p99.

Thirteen novel contributions are presented, including: a unified open-source multi-tool SAST orchestration platform with RAG-backed AI enrichment; a label-free five-signal confidence scorer; an N-gram entropy hallucination filter; a NIST PQC-aware CBoM classifying 62 algorithms; an AST call-graph reachability engine (0\% FP on fixtures); intra-procedural taint analysis (100\% recall on 22 fixtures); the \textbf{P4 Confirmed Tier} as the first open-source SAST stratum to clear the $\geq$80\% commercial autopilot threshold via an external, non-tautological Bandit confidence signal; a \textbf{Verified Remediation Engine} implementing a fully-signed detect~$\to$~exploit~$\to$~patch~$\to$~re-exploit pipeline (13 vulnerability categories) with dual ECDSA-P256 + Dilithium3 attestation of the remediation bundle; and a four-track novel empirical battery (X1--X4) covering generalisation, exploitability, AI-generated-code coverage, and predictive risk.

\vspace{0.8cm}

\noindent\textbf{Keywords:} Static Analysis, Automated Code Review, Security Vulnerability Detection, Retrieval-Augmented Generation, Large Language Models, Confidence Scoring, Call-Graph Reachability, Taint Analysis, Cryptographic Bill of Materials, Post-Quantum Cryptography, DevSecOps, CI/CD Integration, OWASP.

\clearpage
