%% ============================================================
%%  ACKNOWLEDGEMENTS
%% ============================================================
\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
\thispagestyle{plain}

All praise and gratitude are due to Allah, the Most Gracious, the Most Merciful, for granting me the strength, patience, and knowledge to complete this work.

I would like to express my sincere gratitude to my project supervisor, \textbf{Dr.\ Samy AbdelNabi}, for their invaluable guidance, constructive feedback, and continuous encouragement throughout every phase of this project. Their insights on software engineering, artificial intelligence, and academic writing have significantly shaped the quality of this work.

I am deeply grateful to the \textbf{Faculty of Engineering at King Salman International University} for providing an environment that fosters innovation and academic excellence. The faculty members of the Computer Science \& Artificial Intelligence Department have been instrumental in building the foundational knowledge that made this project possible.

Special thanks go to the open-source community behind the static analysis tools integrated in this system --- including Semgrep, Bandit, Safety, PyFlakes, ESLint-security, npm-audit, Gosec, and govulncheck --- whose efforts in building and maintaining these tools made the core of \ACRQA{} feasible.

I am grateful to the creators and maintainers of the Groq inference API and the LLaMA language model family, whose open-access approach to large language model inference enabled the AI-assisted components of this system to operate at zero recurring cost.

My heartfelt thanks go to my family for their unwavering support and patience throughout the long hours of development, testing, and writing. Their encouragement has been a constant source of motivation.

Finally, I thank my colleagues in the Computer Science \& Artificial Intelligence programme at KSIU for the stimulating discussions, peer reviews, and collaborative spirit that enriched this academic journey.

\vspace{1cm}
\begin{flushright}
  \textit{Ahmed Mahmoud Abbas}\\
  \textit{El-Tor, South Sinai, Egypt}\\
  \textit{May 2026}
\end{flushright}

\clearpage
